Cluster Difference Imaging Photometric Survey. III. Rotation and Lithium in an Open Cluster Spanning 500\,Parsecs

### 2021-04-12 cluster_paper/ms.tex
update appendix 2

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we retrieved all available
CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
stars with at least one sector from a CDIPS light curve.  Each of
these stars is brighter than the CDIPS magnitude limit of
$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
source list had $Rp<16$.  The difference (65 stars) is due to 35 stars
from \citet{meingast_2021} which were not available at the time of the
CDIPS reductions, and 30 stars falling on the TESS chip gaps.  The
$Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
roughly to $\bpmrp$ of 2.2, or a spectral type of $\approx$M2V, at the
distance of \cn.

@@

\software{
  %\texttt{arviz} \citep{arviz_2019},
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  %\texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  %dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
  % 	exoplanet:theano},
	\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  \texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  %\texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
	%\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}
\

@@

\section{Detrending details}
\label{app:detrending}

In ``detrending'' for our general variability search, our goal was to
preserve astrophysical variability, while removing systematic
variability.  One particular concern for the TESS light curves is
systematic variability at the timescale of the 14-day satellite orbit,
mostly induced by scattered light from the Earth and Moon.

We therefore turned to the principal components (i.e., the
eigenvectors) calculated following the procedure described by
\citet{bouma_cdipsI_2019}. In brief, these vectors are computed using
a set of ``trend stars'' selected from across each CCD according to
ad-hoc heuristics that (hopefully) lead them to be dominated by {\it
systematic} variability (Sec~3.7.2).

The principal component vectors, also referred to as the eigenvectors,
are rank-ordered by the degree of variance that they predict in the
training set (of ``trend stars'').

We then posit that any given target star's light curve is described as
a linear combination of the eigenvectors.  Optionally, we also
considered the inclusion of additional systematic vectors that could
affect the light curve, such as the CCD temperature, the flux level
measured in the background annulus, and the centroid positions of the
stars on the CCDs.  These can be treated as additional ``features'' in
the linear model.

To determine the coefficients of the linear
model after the full set of eigenvectors (plus optionally
``sytematic'' vectors) had been asssembled,  we explored two possible
methods: ordinary
least squares, and ridge regression. Ridge regression is the same as
ordinary least squares, except it includes an L2 norm with a regularization
coefficient. The regularization coefficient that best applied for any
given target light curve was solved for using a cross-validation grid
search, using \texttt{sklearn.linear\_modelRidgeCV} (CITE).

Each target light curve was mean-subtracted and normalized by its
standard deviation, as were the eigenvectors. The linear problem was
then solved numerically, and the light curve was reconstructed by re-adding the
original mean, and re-multiplying by the standard deviation to ensure
that the variance of the light curve did not change.

We found that the choice of using ordinary
least squares versus ridge regression did not seem to significantly
affect the resulting light curves. In other words, the inclusion (or
lack thereof) of a regularization term did not strongly alter the
best-fitting coefficients.
In the spirit of ``KISS'', we opted to use ordinary least squares.

A few other choices seemed to be more important:

  \item {\it To smooth, or to not smooth the eigenvectors}.
  Ideally, the eigenvectors should be smooth in time. They should not contain
    residuals from {\it e.g.}, eclipsing binaries that snuck their way
    into the template set, and they should also not be intrinsically noisier
    than the target star. If either of these is the case (and we found that
    it sometimes was), it can
    induce extra variability into the PCA ``detrended'' light curves.
    To address this problem, we opted to smooth the eigenvectors using
    a time-windowed filter (with a "biweight" weight scheme, implemented in
    \texttt{wotan} by \citet{hippke_wotan_2019}; window length 1 day, cval 6).
  One issue with this is that systematic sharp features (captured e.g., in
  "spike vectors") no longer are captured, so they end up in the "PCA
  detrended" light curves. They can then be filtered out
  ({\it e.g.}, using rolling outlier rejection), and we prefer this approach to having
  systematic features being {\it injected by} the PCA detrending.

  \item {\it How many eigenvectors to use}.
    A larger number always leads to greater whitening.  In
    \citet{bouma_cdipsI_2019}, we performed a Factor Analysis
    cross-validation to determine the number of eigenvectors to use.
    The typical number adopted based on this analysis was 10--15.
    While this approach should in theory prevent over-fitting, in
    our experience, for stellar rotation it still often lead to
    distorts the signals, especially for rotation signals with small amplitudes
    and periodicities of
    $\gtrsim 3$ days.  (Shorter signals typically are not distorted,
    since the eigenvectors do not contain the high-frequency
    content that leads to the distortions).  For the present analysis, we
    therefore impose the maximum number of eigenvectors to be 5.

  \item {\it Which supplementary systematics vectors to use}.
    We considered using the \texttt{BGV}, \texttt{CCDTEMP},
    \texttt{XIC}, \texttt{YIC}, and \texttt{BGV} vectors, packaged
    with the CDIPS light curves. We found that the background value
    measured in an annulus centered on the aperture, \texttt{BGV},
    tended to produce the best independent information from the PCA
    eigenvectors, and so we adopted it as our only ``supplementary''
    trend vector.  We opted to not smooth it (in hopes that it would
    provide direct complement to the smoothed PCA vectors; 1 sharp
    vector containing literally the background information, plus 5
    smooth vectors).
\end{itemize}

For every ``target star'', we then decorrelated the raw
(image-subtracted and background-subtracted) light curve
using a linear model with ordinary least squares.

>>>>

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we retrieved all available
CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
stars with at least one sector from a CDIPS light curve.  Each of
these stars is brighter than the CDIPS magnitude limit of
$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
source list had $Rp<16$.  The difference (65 stars) is due to 35 stars
from \citet{meingast_2021} which were not available at the time of the
CDIPS reductions, and 30 stars falling on the TESS chip gaps.  At the
distance of \cn, the $Rp=16$\,mag cutoff imposed during the CDIPS
processing corresponds roughly to $\bpmrp$ of 2.2, or a spectral type
of $\approx$M2V.

@@

\software{
  %\texttt{arviz} \citep{arviz_2019},
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  %\texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  %dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
  % 	exoplanet:theano},
	\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  \texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  %\texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
	%\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}
\

@@

\section{Considerations for Removing Systemic Variability from TESS Light Curves}
\label{app:detrending}

In ``detrending'' before searching for stellar rotation signals in the
TESS, our goal was to preserve astrophysical variability, while
removing systematic variability.  One particular concern for the TESS
light curves is systematic variability at the timescale of the 14-day
satellite orbit, mostly induced by scattered light from the Earth and
Moon.

We opted to use a variant of the principal component analysis approach 
described by
\citet{bouma_cdipsI_2019}. This
principal component analysis focuses on a set of ``trend stars'' selected from
across each CCD according using ad-hoc heuristics that on average lead
the light curves to be dominated by systematic variability.  The
resulting principal component vectors, also referred to as the
eigenvectors, are rank-ordered by the degree of variance that they
predict in the training set of trend stars.

We then posit that any given target star's light curve is described as
a linear combination of the eigenvectors.  We also considered the
inclusion of additional systematic vectors that could affect the light
curve, discussed below.  These can be treated as additional ``features'' in the linear
model.  To determine the coefficients of the linear model after the
full set of eigenvectors (plus optional ``sytematic'' vectors) had
been asssembled,  we explored two possible methods: ordinary least
squares, and ridge regression. Ridge regression is the same as
ordinary least squares, except it includes an $\ell^2$ norm with a
regularization coefficient. The regularization coefficient that best
applied for any given target light curve was determined using a
cross-validation grid search, through
\texttt{sklearn.linear\_modelRidgeCV} \citep{scikit-learn}.
Each target light curve was mean-subtracted and normalized by its
standard deviation, as were the eigenvectors. The linear problem was
then solved numerically, and the light curve was reconstructed by
re-adding the original mean, and re-multiplying by the standard
deviation to ensure that the variance of the light curve did not
change.

We found that the choice of using ordinary least squares versus ridge
regression did not seem to significantly affect the resulting light
curves. In other words, the inclusion (or lack thereof) of a
regularization term did not strongly alter the best-fitting
coefficients.  In the spirit of keeping it simple, we opted to use
ordinary least squares.  A few other choices seemed to be more
important:

\begin{itemize}
  %
  \item {\it To smooth, or to not smooth the eigenvectors}.
    If the systematic trends are smooth in time, the eigenvectors should be as well. They should not
    contain residuals from {\it e.g.}, eclipsing binaries 
    in the template set, and they should also not be
    intrinsically noisier than the target star. If either of these is
    the case, extra
    variability can be introduced into the ``detrended'' light curves.  To address
    this problem, we opted to smooth the eigenvectors using a
    sliding time-windowed filter (with a "biweight" weight scheme, implemented
    in \texttt{wotan} by \citealt{hippke_wotan_2019}).
    %; window length 1 day, cval 6).
    One issue with this is that systematic sharp features (captured
    e.g., in ``spike vectors'') are no longer captured, so they are present
    in the detrended light curves. Since they can be filtered out in postprocessing
    ({\it e.g.}, using rolling outlier rejection), we prefer this
    approach to having systematic features being injected by the
    PCA detrending.
  %
  \item {\it How many eigenvectors to use}.
    A larger number always leads to greater whitening.  In
    \citet{bouma_cdipsI_2019}, we performed a factor analysis
    cross-validation to determine the number of eigenvectors to use.
    The typical number adopted based on this analysis was 10 to 15.
    While this approach should in theory prevent over-fitting, in our
    experience, for stellar rotation it still often lead to distorts
    the signals, especially for rotation signals with small amplitudes
    and periodicities of $\gtrsim 3$ days.  Shorter signals typically
    are not distorted, since the eigenvectors do not contain the
    high-frequency content that leads to the distortions.  For the
    present analysis, we therefore set the maximum number of
    eigenvectors to be 5.
%
  \item {\it Which supplementary systematics vectors to use}.  We
    considered using the \texttt{BGV}, \texttt{CCDTEMP}, \texttt{XIC},
    \texttt{YIC}, and \texttt{BGV} vectors, packaged with the CDIPS
    light curves. We found that the background value measured in an
    annulus centered on the aperture, \texttt{BGV}, tended to produce
    the best independent information from the PCA eigenvectors, and so
    we adopted it as our only ``supplementary'' trend vector.  We
    opted to not smooth it, assuming that it would provide direct
    complement to the smoothed PCA vectors.
\end{itemize}

After all these considerations, for every ``target star'', we
ultimately decorrelated the raw (image-subtracted and
background-subtracted) light curve using a linear model with ordinary
least squares.  The vectors considered in the model were the 1
unsmoothed background flux vector, and 5 smoothed PCA eigenvectors.
